Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2015-01-12 parts/def-ph.tex
revised end of section 5

We give in this section the definition and the semantics of the Asynchronous Automata Networks (AAN).
It is a restriction of the classical (synchronous) Automata Networks
where each set of transitions sharing the same label
can only change one local state at a time.
We also discuss how it is related to the Process Hitting framework
(with or without classes of priorities).
In \pref{sec:encodings}, we will also introduce classes of priorities into these models,
and show that they are equivalent to Asynchronous Discrete Networks.

\subsection{Definition of Asynchronous Automata Networks}
\label{ssec:PH}

>>>>

We give in this section the definition and the semantics of the Asynchronous Automata Networks (AANs).
It is a restriction of the classical (synchronous) Automata Networks
where each set of transitions sharing the same label
can only change one local state at a time.
We also discuss how it is related to the Process Hitting framework
(with or without classes of priorities).
Another definition of AANs introducing classes of priorities
is proposed in \pref{sec:encodings},
and we show that they have the same expressivity as AANs.

\medskip

### 2015-01-12 parts/example.tex
revised end of section 5

We held all these experiments on a laptop computer with a 2GHz processor
and 1GB of RAM.
\modMF{
Computations times were in the order of a few hundredths to tenths of a second
in each case (reachability of $\reach_1$ included).
}
To give a comparison, we did the same experiments with a standard symbolic model-checker, libDDD
\cite{libddd}, known for its good performances, the input model being the Boolean network expressed
as a Petri net.
However, due to the large scale of the model, the program runs out of memory for all the experiments.
\modMF{
This limitation shows that studying the dynamics of models is strongly limited to the size of the model for standards model checkers,
while our method is able to conclude with limited memory and computation time usage.
Furthermore, given that the complexity of the static analysis presented in this paper
is polynomial in the size of the models,
it is expected that our work will be scalable to models that are larger,
even by orders of magnitude.
%---~that is, containing thousands of components.
}

\modn{
Finally, we note that similar experiments were conducted in~\cite{PMR12-MSCS}.
However, if these experiments allowed to obtain some information
on this model to some extent,
it did not provide a formal proof of the reachabilities that have been
checked in the example above.
Indeed, the semantics of Process Hitting (without classes of priorities)
do not allow to model accurate Boolean gates,
thus leading to some unwanted behaviours that take the form of temporal shifts.
Only the adding of classes of priorities allows to remove these temporal shifts,
as explained in~\cite{FPMR13-CS2Bio},
with a construction that is equivalent to the AANs presented in \pref{sec:ph}.
Therefore, only the method presented in this paper provides a formal proof that
the observed behaviours are the result of the true dynamics of the system,
which was originally described in a format equivalent to Logical Networks.
}

>>>>

We held all these experiments on a laptop computer with a 2GHz processor
and 1GB of RAM.
Computations times were in the order of a few hundredths to tenths of a second
in each case (reachability of $\reach_1$ included).
To give a comparison, we did the same experiments with a standard symbolic model-checker, libDDD
\cite{libddd}, known for its good performances, the input model being the Boolean network expressed
as a Petri net.
However, due to the large scale of the model, the program runs out of memory for all the experiments,
\modn{%
showing that this method is not scalable.
Because standard model checkers are often PSPACE-complete~\cite{Harel02},
using them to study the dynamics of a model is strongly limited to its size.
On the other hand,
our method is able to conclude with limited memory and computation time usage.
Indeed, the under-approximation developed in this paper comes in two versions:
\begin{itemize}
  \item One version is polynomial in the size of the model,
    and consists in building the saturated graph of local causality once
    and checking \pref{th:approxinf} on this graph only.
    This version is scalable by nature by its low complexity;
  \item The other one is exponential in the size of the model,
    and consists in checking \pref{th:approxinf}
    on every sub-graph obtained by considering sub-sets of the nodes in $\Sol$
    of the saturated graph of local causality.
    Although not as scalable, it is still efficient on this example,
    due to an adapted enumeration of these sub-sets.%
\end{itemize}
It is therefore expected that our work will be scalable to models that are larger,
even by orders of magnitude.
}%
%---~that is, containing thousands of components.

\modMF{
Finally, we note that similar experiments were conducted in~\cite{PMR12-MSCS}.
However, if these experiments allowed to obtain some information
on this model to some extent,
it did not provide a formal proof of the reachabilities that have been
checked in the example above.
Indeed, the semantics of Process Hitting (without classes of priorities)
do not allow to model accurate Boolean gates,
thus leading to some unwanted behaviours that take the form of temporal shifts.
Only the adding of classes of priorities allows to remove these temporal shifts,
as explained in~\cite{FPMR13-CS2Bio},
with a construction that is equivalent to the AANs presented in \pref{sec:ph}.
Therefore, only the method presented in this paper provides a formal proof that
the observed behaviours are the result of the true dynamics of the system,
which was originally described in a format equivalent to Logical Networks.
}
